### 2013-01-29 analysis/SandboxMax/Test.Rmd
Updates after 1/29 meeting with Aaron.

Max's First Analysis
========================================================

This is an R Markdown document. Markdown is a simple formatting syntax for authoring web pages (click the **MD** toolbar button for help on Markdown).

When you click the **Knit HTML** button a web page will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r, cache=TRUE}
getwd()

library (foreign)
workingdata <- read.dbf("../../../Data//R2GIS/CleanData/Sales20052007.dbf")
names(workingdata)
summary(workingdata)

###Preparation for regression analysis, load packages for tests

require(zoo)
require(lmtest)
require(car)

```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
hist(workingdata$SALE_VALUE)
```

Here, let's go ahead and run a regression.

```{r}
model.SaleValue1 <- lm (SALE_VALUE ~ COUNTY_ID + CITY + SALE_YR + SDNUM + GARAGE + ACRES_POLY+HOMESTEAD+FIN_SQ_FT+YEAR_BUILT+MAX+PARK_dist+LAKE_dist+MCA3+MCA5+SHOP_dist+CBD_dist, data=workingdata)
summary(model.SaleValue1)
```